---
title: "Homework 27"
format: html
editor: visual
---

Read through the code in this assignment. You will begin contributing to the code on line 36. Your tasks will be to:

-   Create a dataframe with just the complete cases

-   Review the 1-way ANOVA we ran during class with the full data set

-   Conduct a 2nd 1-way ANOVA using just the data set with complete cases and compare the results to the first analysis.

-   Conduct a linear regression to assess the relationship b/w Growth_rate_mm and Dried_mass_g, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

## Set up our workspace and read in the data

```{r}
rm(list = ls())
library(tidyverse)
library(ggfortify)
library(here)
```

Read in the data.

```{r}
df <- read.csv(here("Data/learn-about-NA.csv"))
str(df)
```

## Dealing with NAs in data analysis

Deciding how to treat NA values in your data can have critical impacts on your data analysis.

Let's look at an example.

Start by creating a data frame from df called df_All_NA, which includes the entire dataset including all NA values.

```{r}
df_All_NA <- df ##Shouldn't need to change anything
```

Now use one of the methods we discussed in class on 13 November to create a second dataframe called `df_No_NA` that is just the complete cases of df.

```{r}
df_No_NA <- df |> drop_na() ##run this command to remove all rows with at least 1 na
```

What if we wanted to test the hypothesis that there is no effect of fertilizer treatment on plant growth rate?

In class on Wednesday 13 November we ran a 1-way ANOVA on the complete data set. Let's review it here:

### First ANOVA using all the data

#### Plot and Guess

```{r}
ggplot(df_All_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot()+
  geom_point()+
  theme_bw()
```

We decided that there was likely an effect of treatment on growth rate, but we weren't sure if both levels of the fertilizer treatment would differ from the control.

#### Create a model:

```{r}
grow.mod<-lm(Growth_rate_mm ~ Treatment, data = df_All_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod)
```

Looks pretty good.

#### Interpret results:

```{r}
anova(grow.mod)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

We have three groups but are fixing one so two are able to vary.

Answer: Because there are 3 levels of our Treatment variable: Control, Level 1, and Level 2. Number of groups - 1 = degrees of freedom therefore 3 - 1 = 2.

Why are there 46 DF for the Residuals? Remember that we determined that there were 49 values in the `Growth_rate_mm` variable using `sum(!is.na(variable))` like this:

```{r}
sum(!is.na(df$Growth_rate_mm))
```

So **there are only 49 data points used in the analysis**, even though the dataframe is 60 rows long!!!

49 - 3 = 46 = degrees of freedom for the residuals.

```{r}
summary(grow.mod)
```

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

No, adding 1 dose of fertilizer doesn't change the growth rate significantly from the control.

Does adding 2 doses of fertilizer change the growth rate significantly from the Control?

Yes, adding 2 doses of fertilizer doesn change the growth rate significantly from the control.

We won't worry about re-plotting our data right now.

### Second Attempt: Complete Cases

Your turn. Run the 1-way ANOVA again, this time with only the dataset that has complete cases and that we called `df_No_NA`

#### Plot and Guess

```{r}
ggplot(df_No_NA, aes(Treatment, Growth_rate_mm))+
  geom_boxplot() +
  geom_point() +
  theme_bw()
```

What do you think?
I would assume that both level 1 and 2 willl be significantly different from the control. I don't believe they will be different from one another.

#### Create a model:

```{r}
grow.mod.no.NA <- lm(Growth_rate_mm ~ Treatment, data = df_No_NA)
```

#### Check our assumptions:

```{r}
autoplot(grow.mod.no.NA)
```

What do you think? Same or different from our first 1-way ANOVA?

The residuals vs fitted plot is a bit different but otherwise looks mostly the same.
#### Interpret results:

```{r}
anova(grow.mod.no.NA)
summary(grow.mod.no.NA)
```

Practice your interpretation.

Why are there 2 DF for the Treatment effect?

We have three groups and are fixing one of them, so since only two can vary our df = 2

Why are there 33 DF for the Residuals?
there are 36 cases in the growth_rate_mm and have 2 x and 1 y variable so our df = 33.

How would you write a results statement?

```{r}
summary(grow.mod.no.NA)
```
We reject the null hypothesis that applying the fertilizer has no effect on the growth rate of plants as p < 0.05 for Treatment level 2.

Practice your interpretation:

Does adding 1 dose of fertilizer change the growth rate significantly from the Control?

Adding 1 dose of fertilizer doesn't change the growth rate significantly from the control.

Does adding 2 doses of fertilizer change the growth rate signficantly from the Control?

Adding 2 doses of fertilizer does change the growth rate significantly from the control.
We won't worry about re-plotting our data right now.

## Linear Regression

Now conduct a linear regression to examine the relationship between `Growth_rate_mm` and `Dried_mass_g`, first with the complete data set and second with the complete cases data set and compare the results of the 2 analyses.

### First attempt: Full data set including NAs (use `df_All_NA`)

You know what to do here. Go for it.

#### Plot and guess

```{r}
ggplot(df_All_NA, aes(Dried_mass_g, Growth_rate_mm))+
  geom_point() +
  theme_bw()
```
I don't expect to see a correlation between dried_mass_g and growth_rate_mm
#### Create model

```{r}
grow.mod.all.NA <- lm(Dried_mass_g ~ Growth_rate_mm, data = df_All_NA)
```

#### Check assumptions

```{r}
autoplot(grow.mod.all.NA)
```

Don't worry if assumptions don't look perfect - focus on normality of the data.
I'm glad you wrote this as my first thought was that my residuals vs leverage looked incredibly weird.
#### Interpret model

Use `anova()` first before using `summary()`

```{r}
anova(grow.mod.all.NA)
```

Why is there 1 degree of freedom for Growth_rate_mm and why are there 34 degrees of freedom for the Residuals?

This time we only have 2 groups and are fixing 1, so we have 1 df.
There are 36 non-NA variables in growth rate and we have 1 x and 1 y variable so our df is 34.

```{r}
summary(grow.mod.all.NA)
```
We reject the null hypothesis that there is no correlation between the Dried_mass_g and Growth_rate_mm.

### Second attempt: Limited data set with only complete cases (use `df_No_NA`)

#### Plot and guess

```{r}
ggplot(df_No_NA, aes(Growth_rate_mm, Dried_mass_g))+
  geom_point() +
  theme_bw()
```
I once again see no correlation between Dried_mass_g and growth rate.

#### Create model

```{r}
grow.mod.dried.no <- lm(Dried_mass_g ~ Growth_rate_mm, data = df_No_NA)
```

#### Check assumptions

```{r}
autoplot(grow.mod.dried.no)
```

Again, don't worry if the assumptions aren't perfect. How do they compare to the first analysis?
Looks about the same as the last one although the normality looks better now.

#### Interpret model

Use `anova()` first then `summary()`.

```{r}
anova(grow.mod.dried.no)
summary(grow.mod.dried.no)
```

Answer the following questions:

1.  How many degrees of freedom are there for Growth_rate_mm? Why?
There is only 1 degree of freedom as there are two variables and 1 is fixed so only 1 can vary.
2.  How many degrees of freedom are there for Residuals? Why?
There are 34 residuals as there are 36 cases in the dataset and we have 1 x and 1 y variable which leaves 34 degrees of freedom.
3.  How are the degrees of freedom in this second model different than the first? What does the answer tell you about the role of NA values in the regression analysis?
The degrees of freedom are the same so I would assume that NA values play little to no role in a regressional analysis as they don't appear to have greatly altered the df.