---
title: "Homework 25"
author: "Andrew Bazuro"
format: html
editor: visual
---
pg 118 to 129
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
daphnia = read.csv(here("data", "Daphniagrowth.csv"))
```

three variables, read in parasite as a factor
```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```

Start by making a boxplot of the data

```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw()
```

Note the text coming in all mashed here, we can fix this using a coord_flip() which alternates the axes.
```{r}
ggplot(daphnia, aes(x = parasite, y = growth.rate)) +
  geom_boxplot() +
  theme_bw() +
  coord_flip()
```

substantial variation in the growth rate between the treatments. highest is 1.2 mm/day.

construct a linear model for a one-way anova.

```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```

```{r}
autoplot(model_grow, smooth.colour = NA)
```

Some concern over normality in the upper right graph (doesn't exactly follow the line).

summary table and anova function
```{r}
anova(model_grow)
```

null hypothesis for a one way anova is that all of the groups come from populations with the same mean. (F-value quantifies how likely this is to be true.)

```{r}
summary(model_grow)
```
control is labelled intercept here. R also deaults to presenting these alphabetically.
```{r}
sumDat <- daphnia |> group_by(parasite) |> summarise(meanGR = mean(growth.rate))
sumDat
```
If ggfortify is not availible there are some ways to find older versions (by googling 'ggfortify cran') or by installing the development version using a series of code (I won't repeat it here as I don't want to install the version but they are on page 130)