---
title: "Homework 15"
author: "Andrew Bazuro"
format: html
editor: visual
---

Today I am reading through Chapter 3 of Getting Started with R and will work through all the material within here. I hope to be comfortable using the following commands: `select()`, `slice()`, `filter()`, `arrange()`, `mutate()`, `with()`, `group_by()`, `summarise()`, `mean()`, and `sd()`

```{r}
rm(list = ls())
library(here)
library(tidyverse)
library(dplyr)
library(ggplot2) #it says ggplot in text but when I attempted to install ggplot I got the error message "package ‘ggplot’ is not available for this version of R". 
compensation = read.csv(here("data", "compensation.csv"))
```

3.1

`summary()` Gives the mean, median, min, and max. As well as levels and sample size for each level of categorical variable.

```{r}
glimpse(compensation)
summary(compensation)
```

3.2

`select()`, `slice()`, `filter()`, `arrange()`, `mutate()` data frame always comes first

3.3

```{r}
select(compensation, Fruit)
#returns a data frame if you use <-
select(compensation, -Root)
```

3.3.2

```{r}
slice(compensation, 2)
slice(compensation, 2:10)
slice(compensation, c(2, 3, 10))
#returns new row numbers in the new df
```

3.3.3

```{r}
with(compensation, Fruit > 80)
#returns what values have > 80 in true/false form, will work well with summarize and group_by to determine statistics
filter(compensation, Fruit > 80)
filter(compensation, Fruit > 80 | Fruit < 20)
```

3.3.4

```{r}
lo_hi_fruit <- filter(compensation, Fruit > 80 | Fruit < 20)
lo_hi_fruit
```

3.4.1

```{r}
head(compensation)
compensation <- mutate(compensation, logFruit = log(Fruit))
```

3.5

```{r}
arrange(compensation, Fruit)
```

3.6

```{r}
select(filter(compensation, Fruit > 80))
compensation |> filter(Fruit > 80) %>% select(Root) # %>% is juts a pipe. (I'm probably going to code these using |> as that what I'm used to if thats ok.)
```

3.7.2

```{r}
summarise(
  group_by(compensation, Grazing),
  meanFruit = mean(Fruit)
)

mean.fruit <- summarise(
  group_by(compensation, Grazing),
  meanFruit = mean(Fruit)
)
```

3.7.3

```{r}
compensation |> group_by(Grazing) |> 
  summarise(meanFruit = mean(Fruit))
```

3.7.4

```{r}
compensation |> 
  group_by(Grazing) |> 
  summarise(
    meanFruit = mean(Fruit),
    sdFruit = sd(Fruit)
  )
```
