---
title: "Homework-26"
author: "Andrew Bazuro"
format: html
editor: visual
---
```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
library(multcomp)
```

```{r}
pea <- read.csv(here("data", "s_r_pea_sections.csv"))
pea$Treatment <- as.factor(pea$Treatment)
```
Plot the data
```{r}
ggplot(pea, aes(x = Treatment, y = pea_section_length_ou)) +
  geom_boxplot() +
  geom_jitter(width = 0.1) +
  theme_bw() +
  coord_flip()
```
Mean for the control is far higher than the rest at 68, otherwise, the order is as follows: 2 % sucrose has a mean of around 64, 2 % glucose is 59, 2 % fructose is 56, while 1 % fructose and 1 % gluxcose is at 56 as well. There are no visible outliers within the data. Based on this I would assume there is a significant difference between the control and all other samples and between the 2 percent sucrose and all other samples, but that there will be no significant difference between the remaining three.

```{r}
ggplot(pea, aes(x = pea_section_length_ou)) +
  geom_histogram() +
  facet_wrap(~Treatment) +
  theme_bw()
```

Create a model
```{r}
model_pea <- lm(pea_section_length_ou ~ Treatment, data = pea)
```

autoplot using ggfortify

```{r}
autoplot(model_pea, smooth.colour = NA)
```
Normality looks good here, as does equal variance.

Run anova and summary table functions

```{r}
anova(model_pea)
```
due to our large F value (49.368) and p value of less than 0.5 we reject the null hypothesis that the mean pea_section_length is the same for each treatment.

```{r}
summary(model_pea)
```
Get the mean pea length
```{r}
sumPea <- pea |> group_by(Treatment) |> summarise(MeanPL = mean(pea_section_length_ou))
sumPea
```
Based on this the mean growth pea length is 12.1 for 1 percent fructose 1 percent glucose, 11.9 for 2 percent fructose, 10.8 for 2 percent glucose, and 6 for 2 percent sucrose

Conducting a TukeyHSD
```{r}
phc1 <- glht(model_pea, linfct = mcp(Treatment = "Tukey"))
summary(phc1)
```

```{r}
cld(phc1)
```

Here we can see there are significant differences between all the control and all other samples, as well as between 2 percent sucrose and all other samples. There was no significant difference between 1 percent fructose 1 percent glucose, 2 percent fructose, and 2 percent glucose.