---
title: "Homework 16"
author: "Andrew Bazuro"
format: html
editor: visual
---

```{r}
rm(list = ls())
library(here)
library(tidyverse)
```

```{r}
cover <- read.csv(here("data", "donnerville_percent_cover.csv"))
## examine dataset using two functions
glimpse(cover)
summary(cover)
```

Rename Site as Location to better reflect what the variable represents

```{r}
cover <- cover |> mutate(Location = Site) |> select(-Site)
```


Set Site, Plot.type, and cover_type as factor variables

```{r}
cover$Location <- as.factor(cover$Location)
cover$Plot.type <- as.factor(cover$Plot.type)
cover$cover_type <- as.factor(cover$cover_type)
```



Set Date as a date variable type
```{r}
cover = cover |> mutate(Date = dmy(Date))
```

Error Checking for Location variable so that areas are consistent.

```{r}
cover$Location[4]
levels(cover$Location)

fix <- which(cover$Location %in% c("donnerville 1", "Donnerville 1", "DONNERVILLE - 1"))
fix2 <- which(cover$Location %in% c("donnerville 2", "Donnerville 2"))
fix3 <- which(cover$Location %in% c("donnerville-3", "Donnerville 3"))

cover$Location[fix] <- "Donnerville 1"
cover$Location[fix2] <- "Donnerville 2"
cover$Location[fix3] <- "Donnerville 3"

cover$Location <- factor(cover$Location)
levels(cover$Location)
```

Error Checking for Plot.type and cover_type

```{r}
levels(cover$Plot.type)
levels(cover$cover_type)
fix4 <- which(cover$cover_type %in% c("bare", "Bare"))
fix5 <- which(cover$Location %in% c("bryophtes", "Bryophytes"))
fix6 <- which(cover$Location %in% c("Vascular", "vasc "))

cover$cover_type[fix] <- "Bare"
cover$cover_type[fix2] <- "Bryophytes"
cover$cover_type[fix3] <- "Vascular"

cover$cover_type <- factor(cover$cover_type)
levels(cover$cover_type)
```

Determine average percent cover by cover type for Donnerville 1

```{r}
cover_Q1 <- cover |> filter(Location == "Donnerville 1") |> select(-Date, -Plot.type, -Slope, -Aspect, -Plot)
##Bare is the last section left so no more work is necessary, but I will select out unnecessary information for ease of reading.
summary(cover_Q1)
```
The average percent cover by cover type for Donnerville 1 is 33.33%.

Determine the average percent vascular plant cover across all sites

```{r}
cover_Q2 <- cover |> filter(cover_type == "Vascular") |> select(cover_type, Percent, Location)
summary(cover_Q2)
```

The average percent of vascular plant cover across all sites is also 33.33%